Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-02-27 TEBenchmarkPaper/article.tex
Alter article text to reflect new data

It is important to ensure that there are no duplicate definitions: theory
exploration does not depend on the names of functions, so neither should our
analysis. For example, consider a problem set which includes a commutativity
theorem for a \texttt{plus} function, and an associativity theorem for an
\texttt{add} function, where the definitions of \texttt{plus} and \texttt{add}
are $\alpha$-equivalent. We would expect an MTE system to conjecture
commutativity and associativity for \emph{both} functions, or for \emph{neither}
function, since they're equivalent. Yet a na\"ive precision/recall analysis
would treat commutativity of \texttt{add} and associativity of \texttt{plus} as
\emph{uninteresting}, since they don't appear in the ground truth.

For this reason, duplicates should be removed (e.g. leaving those which appear
first alphabetically), and any references in the theorem statements updated to
use the remaining definition. In the above example, the \texttt{plus} function
would be removed, and the commutativity theorem updated to reference
\texttt{add} instead.

@@

\begin{itemize}
\item The result would be a single data point, which makes it difficult to
  infer performance \emph{in general}.
\item It is impractical to run any existing MTE system on ``large'' inputs,
  containing more than around 10 definitions.
\end{itemize}

To solve both of these problems we instead \emph{sample} a subset of
definitions. Given a sample size, we choose a subset of that many definitions,
and use that as our MTE input. We generate a corresponding ground truth by
selecting those theorems from the corpus which only contain references to
definitions that appear in our sample.

Unfortunately, uniform sampling of definitions gives rise to a lottery: for a
large corpus and a fixed sample size, it becomes highly unlikely that a sample
will contain the particular definitions required by a theorem. This would cause
most samples to have an empty ground truth, and hence $0$ precision and
undefined recall \emph{independent} of what the MTE system produces. This is
clearly undesirable as an evaluation method.

@@

When we form the ground truth for such samples, we're guaranteed to find at
least one applicable theorem from the corpus (the one we chose as our starting
point).

@@

Given a sample of definitions and a corresponding ground truth, the actual
execution of the MTE system proceeds as in prior work. We must translate the
chosen definitions into the particular input format of the tool under study,
then we time the execution, with a timeout after e.g. 1 hour.

@@

We summarise the runtimes by choosing their median, as it is more robust against
long-running outliers, and hence represents performance for a ``typical''
theory.

@@

\begin{figure}
  \centering
  \input{time.pgf}
  \caption{Running times of QuickSpec on theories sampled from TIP. Each point
    is an individual run, with 30 per size except for 1 (14) and 2 (29). Colour
    indicates the total number of conjectures produced by each run, with failing
    processes shown in red. Some points are shifted horizontally to prevent
    overlaps. Standard box plots are included, although many are obscured due to
    quartiles near zero or timeout. Whiskers show points at 1.5 $\times$ the
    inter-quartile range.}
  \label{figure:quickspec_runtimes}
\end{figure}

@@

We applied our sampling methodology to the 269 TIP definitions for sample sizes
between 1 and 20. We chose 30 samples for each size, but due to the requirement
for samples to admit at least one theorem, only 14 unique samples were obtained
for size 1 and 29 for size 2. A timeout of 3 minutes was imposed, since
preliminary experiments showed that most runs either finish before this time, or
still do not finish after an hour. The running times are shown in
figure~\ref{figure:quickspec_runtimes}.

The only errors we encountered were timeouts (shown in red), and these occurred
for all sample sizes. Whilst the median runtimes are mostly in the range of a
few seconds, the variance shows two distinct regions: sizes below 10 are mostly
finished within 30 seconds, whilst sizes 10 and above have large variance; the
whole upper quartile of size 13 timed out. There is also an unexpected downwards
trend towards size 20, and a slight but consistent reduction in minimum runtimes
from sizes 15 to 20 which is likely an artefact of our testing setup.

Most runs finished quickly and generated few conjectures, with the most being
131 conjectures from a sample of size 20.

@@

There is a clear negative trend in precision as the sample size increases,
indicating that a more aggressive post-processing filter may be useful as we
explore larger theories. Some small samples reached full precision, but nothing
above size 10 exceeded 50\%.

Recall shows a slight downward trend but is mostly flat, averaging between 10\%
and 40\%. This is expected, since adding more definitions to a QuickSpec
signature will never reduce the conjectures it will find. Those conjectures
which QuickSpec failed to find include terms which QuickSpec cannot synthesise
(such as conditional equations, inequalities, terms larger than the search depth
and those with anonymous functions) as well as equations which are removed as
``uninteresting'' by the final congruence closure step.

>>>>

It is important to ensure that there are no duplicate definitions: we are only
concerned with the \emph{logical} content of the input, not the more arbitrary
aspects like the names of functions. For example, consider a problem set which
includes a commutativity theorem for a \texttt{plus} function, and an
associativity theorem for an \texttt{add} function, where the definitions of
\texttt{plus} and \texttt{add} are $\alpha$-equivalent. We would expect an MTE
system to conjecture commutativity and associativity for \emph{both} functions,
or for \emph{neither} function, since they're logically equivalent. Yet a
na\"ive precision/recall analysis would treat commutativity of \texttt{add} and
associativity of \texttt{plus} as \emph{uninteresting}, since they don't appear
in the ground truth.

For this reason, duplicates should be removed, and any references in the theorem
statements updated to use the remaining definition (e.g. chosen based on
lexicographic order). In the above example, the \texttt{plus} function would be
removed, and the commutativity theorem updated to reference \texttt{add}
instead.

@@

\begin{itemize}
\item The result would be a single data point, which makes it difficult to
  infer performance \emph{in general}.
\item It is impractical to run any existing MTE system on ``large'' inputs,
  containing more than a few dozen definitions.
\end{itemize}

To solve both of these problems we instead \emph{sample} a subset of
definitions. Given a sample size, we choose a subset of that many definitions,
and provide those definitions as our MTE input. We generate a corresponding
ground truth by selecting those theorems from the corpus which only contain
references to the sampled definitions.

Unfortunately, uniform sampling of definitions gives rise to a lottery: for a
given sample size, increasing the size of the corpus (which, as detailed above,
is advantageous) makes it less likely that a chosen sample will contain all of
the definitions required by a theorem. The majority of such samples would hence
have an empty ground truth, and thus $0$ precision and undefined recall
\emph{independent} of what the MTE system produces! This is clearly undesirable
as an evaluation method.

@@

When we form the ground truth for such samples, we're guaranteed to find at
least one applicable theorem from the corpus (the one we chose as our starting
point), and hence have a meaningful precision and recall.

@@

Given a sample of definitions and a corresponding ground truth, the actual
execution of the MTE system proceeds as in prior work. We must translate the
chosen definitions into the particular input format of the tool under study,
then we time the execution, with a timeout after e.g. 5 minutes.

@@

We summarise the runtimes by choosing their median, as this is more robust
against long-running outliers, and hence represents performance for a
``typical'' theory.

@@

\begin{figure}
  \centering
  \input{time.pgf}
  \caption{Running times of QuickSpec on theories sampled from TIP. Each point
    is an individual run, with 30 per size except for 2 (29). Colour indicates
    the total number of conjectures produced by each run, with failing processes
    shown in red. Some points are shifted horizontally to prevent overlaps.
    Standard box plots are included, with whiskers showing points at
    1.5~$\times$~the inter-quartile range.}
  \label{figure:quickspec_runtimes}
\end{figure}

@@

We applied our sampling methodology to the 182 definitions which appear in TIP
theorems, for even sample sizes up to 20. We chose 30 samples for each size, but
due to the requirement for samples to admit at least one theorem, only 29 unique
samples were obtained for size 2. A timeout of 5 minutes was imposed, since
preliminary experiments showed that most QuickSpec runs either finish before
this time, or still do not finish after an hour. The running times are shown in
figure~\ref{figure:quickspec_runtimes}.

The only errors we encountered were timeouts (shown in red), and these occurred
for all sample sizes. The running times cluster, with many close to zero and
many exceeding the timeout, but with few in between. There is a clear trend
towards longer times and more timeouts as the sample size increases, such that
most samples of size 20 timed out.

Most runs finished quickly and generated few conjectures, with the most being
116 conjectures from a sample of size 16.

@@

There is a clear negative trend in precision as the sample size increases,
indicating that a more aggressive post-processing filter may be useful as we
explore larger theories. Some small samples reached full precision, but nothing
above size 6 exceeded 50\%.

Recall shows a slight downward trend but is mostly flat, averaging between 20\%
and 50\%. This is expected, since adding more definitions to a QuickSpec
signature will never reduce the conjectures it will find. Those conjectures
which QuickSpec failed to find include terms which QuickSpec cannot synthesise
(such as conditional equations, inequalities, terms larger than the search depth
and those with anonymous functions) as well as equations which are removed as
``uninteresting'' by the final congruence closure step.
